---
output: github_document
---

[![Travis build status](https://travis-ci.org/lucymerobinson/CCAMLRGIS.svg?branch=master)](https://travis-ci.org/lucymerobinson/CCAMLRGIS)

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# CCAMLRGIS R package

A package for reading spatial data that is displayed on the CCAMLR online GIS directly from the webpoint end service into R and creating spatial data for plotting and analyses in R or for exporting and displaying on the online GIS 

## Installation

You can install CCAMLRGIS from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("lucymerobinson/CCAMLRGIS",build_vignettes=TRUE)
```

## Vignettes 

Vignettes provide examples of how functions in the package work.
To browse vigenettes you can follow the install instructions above and then use:
browseVignettes("CCAMLRGIS")

Or if you want to browse examples online prior to install these can be accesses at (https://lucymerobinson.github.io/CCAMLRGIS/)

To view the worked examples navigate to the Articles tab at the top of the page and select the "Introduction to the CCAMLRGIS R package"




